\vspace{-0.5cm}The X-shooter GTO sample of GRB afterglow and host Galaxy spectra

### 2016-10-07 document/XSGRB_sample.tex
Update XSGRB_sample.tex
GRB120118B, GRB120119A, GRB120211A

	\section{Reduction scheme}
	\input{tables/burst_overview.tex}

	\section{Results}

	\subsection{Spectral resolution}
	The afterglow spectra described in this paper are obtained in
	Target-of-Opportunity (override) mode.
	In most cases there is therefore little possibility to tweak slit widths to the
	seeing at the time of observations (i.e. to optimise spectral resolution and
	signal to noise), and almost all our data is therefore taken with a fixed set
	of slit widths and binning, described above.
	In a fair number of cases, the seeing full width at half maximum (FWHM) is
	considerably smaller than the silt width, and the delivered spectral resolution
	will then be determined by the seeing rather than slit width, as afterglows are
	point sources (this is evidently not the case for extended sources, e.g for
	host galaxies).
	The delivered resolution for slit width dominated spectra post-reduction and
	extraction can easily be determined from the bright sky emission lines.
	For afterglow spectra with very high signal to noise, the delivered spectral
	resolution can at times be determined from the science data themselves.
	However, in the presence of multiple velocity components in absorption, other
	forms of line broadening, and a lack of lines at some redshifts, this is
	difficult to do at poorer signal to noise ratios (the majority of spectra in
	our sample).
	A broad starting value for the expect resolution will help fitting of these
	spectra, and can be important in upper limit determination, and for this reason
	we construct a aim to construct a crude relation between the seeing and the
	delivered resolution at our slit width, binning, and reduction pipeline
	settings.
	To this end we use observations of telluric standard stars that are taken with
	identical instrument settings as our afterglow spectra, usually just after the
	science data, as part of the ESO X-shooter calibration plan.
	These spectra have been reduced together with the afterglow spectra, using
	identical pipeline settings with the same version of the pipeline.
	First we fit a Gaussian function in the spatial direction of the trace of the
	standard star at 792 nm (i.e. in the VIS arm).
	After this, we fit a series of  20 telluric absorption lines in the telluric
	standard star spectra with Gaussians, taking care to select transitions that
	are not almost-resolved multiples, should be intrinsically unresolved, and are
	in areas with well defined continuum flux.
	We pick 34 telluric standard stars spanning a range of DIMM seeing values, with
	the majority between $0.5-1.5 arcsec$.
	The resulting distribution of spectral FWHM (km/s) as a function of spatial
	FWHM at 792 nm is fairly well described by a linear relation $a + b*x$, with
	$x$ the spatial FWHM in pixels (with 0.15 arc sec per pixel),  $a= 21.4 \pm
	1.3$ km/s, $b=1.4 \pm 0.2$.
	We use this linear relation as a way to estimate the spectral resolution for
	medium to poor signal to noise afterglow spectra in the VIS arm.
	To extend this to the UVB and NIR arm, we measured a series of lines in NIR arm
	spectra of a subset of 19 sources used for the VIS arm above, and find that the
	resulting distribution is consistent with a simple scaling of the VIS arm
	relation by the ratio of resolutions of the NIR and VIS arm for unresolved,
	slit filling, sources as given on the ESO instrument website.
	The UVB arm contains no suitable absorption lines to use, and we therefore use
	a scaled value as in the NIR arm.
	While this simple method is not terribly accurate (for one, the spatial profile
	of the trace is not a perfect Gaussian), but it gives a sufficiently accurate
	estimate for the analysis of these poor signal to noise science spectra.

@@

>>>>

	\section{Reduction scheme}

	Additionally, the wavelength solution for all bursts have been re-calibrated. Using a synthetic sky spectrum \citep{Noll2012, Jones2013}, a first refinement of the wavelength solution have been obtained by cross-correlating with the observed sky.

	Both a multiplicative and an additive offset has been tested, but in terms of $\chi^2$, the model with only a multiplicative offset is preferred.

	\section{Results}
	\input{tables/burst_overview.tex}	
	\subsection{Spectral resolution}
	The afterglow spectra described in this paper are obtained in
	Target-of-Opportunity (override) mode.
	In most cases there is therefore little possibility to tweak slit widths to the
	seeing at the time of observations (i.e. to optimise spectral resolution and
	signal to noise), and almost all our data is therefore taken with a fixed set
	of slit widths and binning, described above.
	In a fair number of cases, the seeing full width at half maximum (FWHM) is
	considerably smaller than the silt width, and the delivered spectral resolution
	will then be determined by the seeing rather than slit width, as afterglows are
	point sources (this is evidently not the case for extended sources, e.g for
	host galaxies).
	The delivered resolution for slit width dominated spectra post-reduction and
	extraction can easily be determined from the bright sky emission lines.
	For afterglow spectra with very high signal to noise, the delivered spectral
	resolution can at times be determined from the science data themselves.
	However, in the presence of multiple velocity components in absorption, other
	forms of line broadening, and a lack of lines at some redshifts, this is
	difficult to do at poorer signal to noise ratios (the majority of spectra in
	our sample).
	A broad starting value for the expect resolution will help fitting of these
	spectra, and can be important in upper limit determination, and for this reason
	we construct a aim to construct a crude relation between the seeing and the
	delivered resolution at our slit width, binning, and reduction pipeline
	settings.
	To this end we use observations of telluric standard stars that are taken with
	identical instrument settings as our afterglow spectra, usually just after the
	science data, as part of the ESO X-shooter calibration plan.
	These spectra have been reduced together with the afterglow spectra, using
	identical pipeline settings with the same version of the pipeline.
	First we fit a Gaussian function in the spatial direction of the trace of the
	standard star at 792 nm (i.e. in the VIS arm).
	After this, we fit a series of  20 telluric absorption lines in the telluric
	standard star spectra with Gaussians, taking care to select transitions that
	are not almost-resolved multiples, should be intrinsically unresolved, and are
	in areas with well defined continuum flux.
	We pick 34 telluric standard stars spanning a range of DIMM seeing values, with
	the majority between $0.5-1.5 arcsec$.
	The resulting distribution of spectral FWHM (km/s) as a function of spatial
	FWHM at 792 nm is fairly well described by a linear relation $a + b*x$, with
	$x$ the spatial FWHM in pixels (with 0.15 arc sec per pixel),  $a= 21.4 \pm
	1.3$ km/s, $b=1.4 \pm 0.2$.
	We use this linear relation as a way to estimate the spectral resolution for
	medium to poor signal to noise afterglow spectra in the VIS arm.
	To extend this to the UVB and NIR arm, we measured a series of lines in NIR arm
	spectra of a subset of 19 sources used for the VIS arm above, and find that the
	resulting distribution is consistent with a simple scaling of the VIS arm
	relation by the ratio of resolutions of the NIR and VIS arm for unresolved,
	slit filling, sources as given on the ESO instrument website.
	The UVB arm contains no suitable absorption lines to use, and we therefore use
	a scaled value as in the NIR arm.
	While this simple method is not terribly accurate (for one, the spatial profile
	of the trace is not a perfect Gaussian), but it gives a sufficiently accurate
	estimate for the analysis of these poor signal to noise science spectra.

@@

	\subsection{GRB120118B (z =	2.943)}
	The data presented here also formed the basis of GCN \#
	14225\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14225.gcn3}}, but is not
	published otherwise. This late-time observation of the host of GRB120118B
	consists of 3.6 ks exposures and contains emission lines belonging to \oii and
	\oiii at $z =	2.943$, suggested to be redshift of the host.

	\subsection{GRB120119A (z =	1.728)}
	The data presented here has not been published before. Three epochs of
	observations have been obtained, the first two immediately after the burst, and
	the last one long after the afterglow had faded. Starting 1.4 hours after the
	\textit{Swift} trigger, the first epoch contains bright afterglow continuum.
	Rich in absorption features belonging to a multitude of ions, $z =	1.728$ is
	estimated for the host with intervening systems at $z =	1.476$, $z = 1.214$, $z
	= 0.662$ and $z = 0.632$. The second epoch, obtained 4.5 hours after the burst
	contains the fading afterglow. A third epoch is obtained $>1$ year after the
	GRB in which emission lines from \hb~and \ha~are found at the redshift of the
	host, confirming the association of the absorption line system and the host.

	\subsection{GRB120211A (z =	2.346)}
	The data presented here has been published in \citet{Kruhler2015}. Two
	observations of the host of GRB120211A has been obtained, starting 2013.02.17,
	$> 1 year$ after the burst has faded. A redshift for this object has been
	reported by \citet{Kruhler2015} and the features seen by those authors are
	reproduced in these reductions, confirming $z =	2.346$.
